---
title: "Fml_assignment_3"
author: "Abhinav"
date: "2024-03-10"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
UN_bank <- read.csv("/Users/keerthanavonteddu/Desktop/abhinav/UniversalBank.csv")
summary(UN_bank)
```


```{r}
library(dplyr)
```

```{r}
library(ISLR)
```

```{r}
library(e1071)
```

```{r}
library(caret)
```

```{r}
library(class)
```

```{r}
library(ggplot2)
```

```{r}
library(tidyr)
```

```{r}
library(gmodels)
```

```{r}
library(lattice)
```


```{r}
UN_bank$Personal.Loan <- factor(UN_bank$Personal.Loan)
UN_bank$Online <- factor(UN_bank$Online)
UN_bank$CreditCard <- factor(UN_bank$CreditCard)
```


# 1.
```{r}
set.seed(231)
Index_train <- createDataPartition(UN_bank$Personal.Loan,p = 0.6,list = FALSE)
Train_df <- UN_bank[Index_train,]
validation.df <- UN_bank[-Index_train,]
```



```{r}
pivot_table <- xtabs(~ CreditCard + Online + Personal.Loan,data = Train_df)
ftable(pivot_table)
```

# 2. 
```{r}
probabality = 48/(48+474)
probabality
```

# 3. 
```{r}
table(Personal.Loan = Train_df$Personal.Loan, Online = Train_df$Online)
table(Personal.Loan = Train_df$Personal.Loan, CreditCard = Train_df$CreditCard)
table(Personal.Loan = Train_df$Personal.Loan)
```

# 4. 
# (i)
```{r}
Probabality.1 <- 90/(90+198)
Probabality.1
```


# (ii)
```{r}
Probabality.2 <- 165/(165+123)
Probabality.2
```

# (iii)
```{r}
Probabality.3 <- 288/(288+2712)
Probabality.3
```

# (iv)
```{r}
Probabality.4 <- 782/(782+1930)
Probabality.4
```

# (v)
```{r}
Probabality.5 <- 1624/(1624+1088)
Probabality.5
```

# (vi)
```{r}
Probability.6 <- 2712/(2712+288)
Probability.6
```


# 5. 
```{r}
calculated_probabilities <- (Probabality.1 * Probabality.2 * Probabality.3)/((Probabality.1 * Probabality.2 * Probabality.3) + (Probabality.4 * Probabality.5 * Probability.6))
calculated_probabilities
```

# 6.
We obtained a #Value of 0.09195402 from question 2, which is nearly identical to the value of 0.09918921 from question 5. The precise approach and the naïve bayes technique are identical in every way, however the former requires the precise categorization of every independent variable for prediction, while the latter does not. The figure obtained from question 2 is more accurate, as we can validate. Considering that the pivot table's precise values were used.


# 7.
```{r}
naviebayes.model <- naiveBayes(Personal.Loan ~ Online + CreditCard, data = Train_df)
to.predict = data.frame(Online=1, CreditCard= 1)
predict(naviebayes.model, to.predict,type = 'raw')
```

#The value we obtained from task 5 is 0.09918921, and the value we get from question 7 is 0.9070972. The outcome is nearly identical to what we obtained from task 5. #The rounding results in a very little change. The output's rank order will remain unaffected by the discrepancy.



























































































